Species richness and food-web structure jointly drive community biomass and its temporal stability in fish communities

### 2021-07-06 manuscript/bef_stability/manuscript.Rmd
rewiew according to reviewer 1 and 2

---
title: Species richness and food-web structure jointly drive community biomass and its temporal stability in fish assemblages 
author: Alain Danet, Maud Mouchet, Willem Bonnaffé, Elisa Thébault, Colin Fontaine
date: \today 
output:
  bookdown::pdf_document2:
    toc: false 
    fig_caption: true 
    keep_tex: true
  word_document:
fontsize: 12pt
header-includes:
   - \usepackage{setspace}
   - \doublespacing
   - \usepackage{lineno}
   - \linenumbers
geometry: margin=2cm
bibliography: references.bib
csl: ecology-letters.csl
---

@@

- Number of words:
  - Abstract: 149
  - Main text: 4930

- Number of references: 92
- Number of figures: 3
- Number of tables: 1
- Number of text boxes: 0

@@

Biodiversity-ecosystem functioning (BEF) and food-web complexity-stability
relationships are central to ecology. However, they remain largely untested in
natural context. Here we quantify the links among environmental conditions,
richness, food-web structure, annual biomass and its temporal stability using a
standardized monitoring dataset of 99 stream fish communities spanning from 1995
to 2018. We first reveal that both richness and average trophic level are
positively related to annual biomass, with effects of similar strength. Second,
we show that mean trophic level decrease population variability thereby
increasing community stability, and contrary to expectation, a destabilizing
effect of species richness. Finally, we find that environmental conditions
affect both biomass and its stability via effects on richness and network
structure. Strikingly, the variation of stability are mediated by population
stability rather than synchrony, which contrasts with results from
single-trophic communities and might be characteristic of multi-trophic
communities.

@@

Species richness has been found to increase primary productivity, community
biomass and its temporal stability in a diversity of ecosystems and taxa both
empirically and experimentally [@tilman_productivity_1996;
@cardinale_species_2002; @tilman_biodiversity_2006; @franssen_annual_2011;
@grace_integrative_2016; @pennekamp_biodiversity_2018; @houlahan_negative_2018;
@olivier_urbanization_2020].  However, these studies mainly focused on
communities composed of a single trophic level, generally primary producers,
thereby ignoring potential effects of food-web structure [but see
@scherber_bottom-up_2010]. Conversely, theoretical works suggest that both
species richness and food-web structure affect ecosystem properties such as
productivity, total biomass and its stability. For example, the presence of
species at higher trophic levels may increase total primary production, through
the selection of primary producers which are larger and more complementary
[@wang_biodiversity_2018]. In communities including producers and consumers, the
connectance which measures the degree of consumer generalism is predicted to
modulate the diversity-productivity relationship [@thebault_food-web_2003;
@thebault_cascading_2007]. Similarly, theoretical studies generally show strong
links between food-web structure and stability [e.g.  @may_will_1972;
@dunne_network_2006; @neutel_reconciling_2007; @duffy_why_2009;
@stouffer_compartmentalization_2011].  Overall, the evidences from theoretical
model suggests that both species richness and food-web structure should affect
community biomass and stability, but empirical evidences of the effects of
network structure on community biomass and stability are lacking.

While community stability can be measured in many ways in theoretical studies,
most empirical studies measure community stability as the inverse of the
coefficient of variation of community productivity, biomass or abundance, also
called temporal stability [@kefi_advancing_2019]. This stability metric is also
well grounded in theory [e.g. @thebault_trophic_2005; @loreau_species_2008;
@thibaut_understanding_2013] and can be decomposed into two components:
population stability and synchrony [@thibaut_understanding_2013].  Studies
considering a single trophic level [@yachi_biodiversity_1999;
@tilman_biodiversity_2006; @loreau_species_2008; @olivier_urbanization_2020] as
well as theoretical prediction for multitrophic communities
[@thebault_trophic_2005] tend to highlight that asynchrony in population
dynamics as key for the stabilizing effects of species richness on community
biomass. Theory also predicts that higher food-web connectance can
counterbalance the positive effect of species richness on the temporal stability
of community biomass, because it may increase interaction strength and
thereby increases population variability, leading to a negative relationship
between species richness and the temporal stability of populations
[@thebault_trophic_2005]. Predators can also have cascading effects on both
population variability and synchrony at lower trophic levels
[@teng_dynamics_2004; @shanafelt_stability_2018]. Overall, these theoretical
studies show that the effects of species richness and food-web structure, such
as connectance and average trophic level, are potentially conflicting or
synergistic in such a way that they need to be addressed jointly.

The joint study of the effects of species richness and food-web structure on total
biomass and its stability is even more important in the context of environmental
changes. Recent studies comparing natural communities have shown that
environmental characteristics determine in part community biomass
[@grace_integrative_2016; @woods_testing_2020] and its temporal stability
[@franssen_annual_2011; @hansen_climate_2013; @bluthgen_land_2016;
@de_boeck_patterns_2017; @friedpetersen_drivers_2020]. The environment can
affect ecosystem function or stability directly. For example, larger freshwater ecosystems
tend to have more available resources and thus community biomass
[@post_ecosystem_2000; @doi_resource_2009] while agricultural intensification
destabilizes bird communities by decreasing population stability
[@olivier_urbanization_2020]. The effects of environment can also go through
effects on species richness and food-web structure, as larger ecosystems host
food-webs with higher diversity and trophic levels [@mchugh_dual_2010] whereas
temperature can affect food-web diversity and structure in contrasted ways
depending on the ecosystem studied [@ogorman_unexpected_2017;
@gauzens_biodiversity_2020]. Environmental characteristics have thus to be
accounted for if we are to tease apart the effects of species richness and
foodweb structure on community biomass and its temporal stability.

---
#### Stream fish communities is a good model to study effects of horizontal & vertical diversity
---

Stream fish communities constitute a good model to study the effects of network
structure on total biomass and its temporal stability because their trophic
interactions are well known,. A substantial body of literature supports that
trophic interactions can be inferred through body size and diet informations
[@jennings_weak_2001; @beckerman_foraging_2006; @gravel_inferring_2013;
@brose_predator_2019; @portalier_mechanics_2019]. Using long-term monitoring of
stream fish communities [@edeline_ecological_2013], we propose one of the first
attempts to test empirically the theoretical expectations on the
diversity-stability relationships according for community, trophic complexity
and environmental characteristics.

@@

Fish communities were monitored across stream sections in France on the period
1995-2018 by the French Office of Water and Aquatic Ecosystems (ONEMA) using
electrofishing. Two different standardized protocols were used for small and
large streams (detailed in appendix B, part 1.1). In the case where a stream was
sampled alternatively with one of the two sampling protocols, we only included
data derived from the most frequently used sampling method. The sampling season
spanned from late spring to autumn.  There was low heterogeneity of the sampling
month within stations, with the median sampling month being mid-August where the
median standard deviation of the sampling month was $0.7$ month, i.e. three weeks.

We selected the stations that had been sampled ten years or more. As the concept
of community stability often refersred to the variability around an equilibrium
point [@donohue_navigating_2016], which is rarely observed or difficult to
characterize in natural communities. Then, we selected the stations that did not
show temporal trends in community biomass (detailed in appendix B, part 2).
However, the results were robust to the relaxation of this hypothesis (Fig. S7).
We selected the stations that had been sampled ten years or more. As the concept
of community stability referred to variability around an equilibrium, which is
rarely observed or difficult to characterize in natural communities, only those
stations that did not show temporal trends in community biomass were selected
(detailed in appendix B, part 2). However, the results were robust to the
relaxation of this hypothesis (Fig. S7). These selection steps resulted in 99
stations distributed over seven hydrographic basins (Fig. \@ref(fig:matmet)A).

@@

The body size of all fish individuals from a given species was not recorded when they were too
numerous (Fig. S1, appendix B, part 1.1). The body size of the unmeasured
individuals was inferred under the assumption that species body size followed a
normal distribution (Fig. S1, appendix B, part 1.2). The body mass of each
individual was then estimated following the results of two meta-analyses on
fishes [@lleonart_removing_2000; @froese_cube_2006], using an allometric law
$B_i = 0.01 \times l_i^{3.03}$, with $B$ the body mass in grams, $l$ the body
size in millimeters and $i$ the individual. Population biomass and community biomass at
each sampling event were then calculated by summing individual body masses. To
account for the variation in sampling effort among sites, population biomass and
community biomass were corrected by the surface sampled (range `r range_surface[1]` -
`r range_surface[2]` square metres), and expressed in grams per square meter
($gm^{-2}$). In the analysis, the community biomass of a site was defined as
the median of the community biomass across sampling years.

@@

Temporal biomass stability was defined as the inverse of biomass variability.
Variability was computed as Coefficient of Variation (CV) of the community biomass over
time ($CV_i = \sigma_i \mu_i^{-1}$, with $\sigma_i$ and $\mu_i$ being
respectively the standard deviation and the temporal average of the annual
community biomass of the station $i$). Then, temporal stability ($S_i$) was
equal to $CV_i^{-1} = \mu_i \sigma_i^{-1}$.

The variability of community biomass ($CV_{com, i}$) can be decomposed 
into a
synchrony ($\phi$) component and a weighted average population variability
($\overline{CV_{sp, i}}$) component according to @thibaut_understanding_2013:

@@

To assess the mechanisms driving stability of community biomass, we thus in addition measure
the synchrony and the weighted average population variability $\overline{CV_{sp,
i}}$, the average CV at the species-level in the station $i$, weighted by the
relative biomass of the species, $B_{k,i}$.

Synchrony of the station $i$, $\phi_i$, is the ratio between the variance of
community biomass over time ($\sigma^2_{x_{T,i}}$) and the sum of the variance of
population biomass over time ($\sigma_{x_{k,i}}$), $i$ being the station $i$ and $k$
the species $k$ [@loreau_species_2008].

@@

Species richness in a given site was defined as the total number of species present across
all the sampling events. We controlled species richness for sampling effort by
dividing the total number of species by the total surface sampled in the given
station. Species richness was then expressed in number of species per square
meter sampled.

### Network inference

@@

We determined the network structure in two steps: (1) build a metaweb describing the
trophic interactions between fishes and the resources (Fig.
SXX), then (2) determine network of a given community by
extracting the metaweb according to the fishes present at a given
site and sampling event (Fig. \@ref(fig:matmet)C and D).

The trophic interaction network was inferred from body size and ontogenic diet
shift following previous studies [@edeline_ecological_2013;
@gravel_inferring_2013; @poisot_structure_2016; @brose_predator_2019]. The diet
of stream fishes, as for many organisms, shifts throughout an individual's life,
and thereby with body size. For example, larvae and small juveniles feed on
plankton while adults may feed on algae and fishes, with bigger individuals
eating bigger fishes. Each fish species was divided in nine body size
classes, each corresponding to a trophic species (appendix B, Fig. S2). A
trophic species constitutes a group of individuals from the same species and
size class, which are expected to share the same sets of prey and predators.
Additionally to the trophic species, we added seven resource nodes present in
the ontogenic food diet database: detritus, biofilm, phytoplankton, zooplankton,
macrophages, phytobenthos and zoobenthos (grey nodes in Fig. \@ref(fig:matmet)C
& D). The trophic species and the resources constituted the 412 nodes of the
metaweb, i.e. 45 fish species divided into nine body size classes plus seven
resource nodes. The inference of trophic interactions was shown to be more
accurate when considering body size classes, i.e. trophic species, than species
[@jennings_weak_2001].

The feeding interactions of a fish node, i.e. a trophic species, were determined
by both its species identity and its body size. The ontogenic fish diet
database was filled according to fishbase and literature (appendix A, Table S2). Each
fish species had two or three life stages which are delimited by its body size
range (appendix B, Fig. S2). The life stage of a fish trophic species was determined by 
the center of its body size range, hereafter midpoint (appendix B, Fig. S2).

For a piscivorous trophic species, the fish-fish trophic links were defined in
two steps and based on the body size ratio between predators and prey. First of
all, the predation window of a piscivorous trophic species was defined as 3% to
45% of its midpoint [@mittelbach_ontogeny_1998; @claessen_impact_2002;
@hart_handbook_2008]. Then, a trophic link was set between the piscivorous
trophic species and every trophic species whose midpoint in bodysize range was
included in the predation window (appendix B, Fig. S2). The fish-resource trophic links were
set between the trophic species and the resources nodes according the food items
of their size class.  Lastly, the trophic links among resource nodes were set
according to the literature [@allan_stream_2007; @hart_handbook_2008].

### Network structure

For each food-web, corresponding to a particular site and year, we computed the
connectance and mean trophic level. Connectance describes the level of
generalism in the network and is calculated as the actual number of interactions
divided by the number of interactions if all species interact ($C = L / N^2$, L
and N being resp. the number of links and of nodes).  The trophic level of the
nodes was computed as one added to the average trophic level of its food item.
The average trophic level $\overline{T}$ was computed as the average trophic
level of the trophic species weighted by their biomass (excluding resource nodes
for which we had no biomass information), so $\overline{T} = \sum_i^N{B_iT_i
\times (\sum_i^N{B_i})^{-1}}$, with $T_i$ and $B_i$ being respectively the
trophic level and the biomass of the node $i$. For each site, the median of
connectance and average trophic level over time was used in subsequent analysis
(Fig. \@ref(fig:matmet)C & D).

@@

We characterised the sites by their altitude, slope, distance from
source, stream width, depth and strahler order, i.e. its position in the
stream network hierarchy. We obtained data on water temperature, flow and Biological Oxygen
Demand (BOD) from the naiades
database
[(www.naiades.eaufrance.fr)](http://www.naiades.eaufrance.fr/). As data from naiades are
not collected on the same site as our fish sampling sites, we performed an
interpolation of the water temperature and flow based on the spatial stream
network models [@isaak_applications_2014; @hoef_ssn:_2014]. The interpolation procedure
is detailed in appendix B (part 1.4).

We had one environmental variable value for each sampling event, except
for water temperature for which the record began in 2006. Then, we computed the
average and temporal CV values of water temperature, flow, BOD, river width and
depth. As habitat and environmental variables are often collinear, a principal
component analysis (PCA) was performed on the 14 scaled variables. We kept in
the analysis the two first principal components based on the elbow method, on
the eigenvalues distribution (appendix B, Fig. SXX). The two first principal components
explained `r tot_var_expl_pca`% of the variance (Fig. \@ref(fig:matmet)B). A varimax axis rotation was
performed to maximize the representation of the environmental variables by the principal
components [@kaiser_varimax_1958; @revelle_psych_2019]. The first axis
represented a gradient from small to big streams, characterized by their width
and depth, but also a gradient from stream close to far from the source. We
reversed the values of the second axis, such as it represented a gradient of
increasing average water temperature, average BOD and decreasing altitude and slope.
Further description of the variables are provided in appendix B (part 1.4).

@@

We analyzed the relationships among environment, species richness, food-web
structure, community biomass and its temporal stability with structural equation
models as they enable the assessment of potential indirect and direct causal
relationships among multiple variables [@grace_structural_2008;
@grace_causal_2014; @lefcheck_piecewisesem_2016; @grace_integrative_2016].  We
defined a structural equation model explaining the median annual community
biomass as a function of the two synthetic environmental variables, species
richness and food-web structure [@maureaud_biodiversityecosystem_2019;
@woods_testing_2020]. We assumed that both species richness and food-web
structure directly affect community biomass (Fig. \@ref(fig:sem-bm)A) and that
species richness affected food-web structure [@winemiller_must_1989;
@dunne_network_2006; @thebault_stability_2010].

We used the same model structure to assess the drivers of temporal stability
of community biomass (Fig.  \@ref(fig:sem-stab)A). Instead of the
annual community biomass, we modelled the synchrony and $\overline{CV_{sp}}$ which
in turn determined the temporal stability of community biomass. We linearized
the mathematical relationship linking temporal stability to $\overline{CV_{sp,
i}}$ and $\phi$ by taking the log of
temporal stability such as

@@

The structural equation models were based on Gaussian linear mixed-effects models. All variables
except the connectance and average trophic level
were log-transformed to fulfill model assumptions (see residual plots and data
transformation in Fig. S5 and S6, Table S5, appendix B). We set the seven
hydrographic basins as a random effect on the intercept to
account for their heterogeneity. We checked for the absence of multicollinearity 
with Variance Inflation Factor (VIF, Table S6, appendix B). The slopes reported
in the main text were standardized in order to compare their magnitude ($r_\delta
= \beta \frac{\sigma_x}{\sigma_y}$, $\beta$ being the unstandardized slope,
$\sigma_x$ and $\sigma_y$ being respectively the standard deviation of the
predictive variable $x$ and of the response variable $y$). Therefore, the
standardized coefficients expressed the variation of $x$ and $y$ in standard
deviation units. According to the equation (\@ref(eq:stab)), the unstandardized
coefficients of synchrony and $\overline{CV_{sp}}$ determining temporal
stability of biomass are respectively $-0.5$ and $1.0$. When standardized, a 
comparaison of those coefficients comes down to a comparaison of the standard
deviation of these two variables. The paths whose slope had a $P$ value inferior
or equal to $0.05$ were reported in the main text, all the path values are
presented in Table S7 and S8.

Indirect effects on community biomass and its temporal stability were computed
by multiplying slopes along the path. Only significant direct slopes were
included in the calculation of indirect effects. Total effects were computed as
the sum of direct and indirect effects.

Environmental variables were interpolated with the `openStars` and `SSN`
R packages [@kattwinkel_openstars_2018; @hoef_ssn:_2014]. Linear models and SEMs
were computed with respectively the `nlme` and `piecewieseSEM` R packages
[@pinheiro_nlme_2020; @lefcheck_piecewisesem_2016]. The code used for the
analysis and to generate the manuscript are accessible on github
(alaindanet/fishcom). The analyses were performed with R version 3.6.3
[@r_core_team_r_2020].

@@

```{r, cache = FALSE}
sem_caption <- paste0(
  "Metamodel of the Structural Equation Models for the (A) median of annual
  community biomass (in grams by square meter) and (C) its temporal stability.",
  "Results of the Structural Equation Models for
  (B) the median of annual community biomass and (D) its temporal stability. Red and black
  arrows represent respectively negative and positive path coefficients. The
  path coefficients are standardized. For convenience, only significant
  paths (i.e. $P < 0.05$) are shown. Rsq: Adjusted $R^2$; Avg:
  Average, CVsp: weighted temporal CV of the population; PCA1 and PCA2: respectively the first and
  second axis of the PCA performed on environmental variables (see methods)."
)
```

>>>>

---
title: Species richness and food-web structure jointly drive community biomass and its temporal stability in fish communities
author: Alain Danet, Maud Mouchet, Willem Bonnaffé, Elisa Thébault, Colin Fontaine
date: \today 
output:
  bookdown::pdf_document2:
    toc: false 
    fig_caption: true 
    keep_tex: true
  word_document:
fontsize: 12pt
header-includes:
   - \usepackage{setspace}
   - \doublespacing
   - \usepackage{lineno}
   - \linenumbers
geometry: margin=2cm
bibliography: references.bib
nocite: |
    @declerck_species_2006
csl: ecology-letters.csl
---

@@

- Number of words:
  - Abstract: 148
  - Main text: 4917

- Number of references: 102
- Number of figures: 3
- Number of tables: 1
- Number of text boxes: 0

@@

Biodiversity-ecosystem functioning (BEF) and food-web complexity-stability
relationships are central to ecology. However, they remain largely untested in
natural contexts. Here we estimated the links among environmental conditions,
richness, food-web structure, annual biomass and its temporal stability using a
standardised monitoring dataset of 99 stream fish communities spanning from 1995
to 2018. We first revealed that both richness and average trophic level are
positively related to annual biomass, with effects of similar strength. Second,
we found that community stability is fostered by mean trophic level, while
contrary to expectation, it is decreased by species richness. Finally, we found
that environmental conditions affect both biomass and its stability mainly via
effects on richness and network structure. Strikingly,
community stability was mediated by the variation of population stability rather
than synchrony, which contrasts with results from single-trophic communities. We
discuss the hypothesis that it could be a characteristic of multi-trophic
communities.

@@

Species richness has been found to increase primary productivity, community
biomass and its temporal stability in various ecosystems and taxa both
experimentally and in natural settings [@tilman_productivity_1996;
@cardinale_species_2002; @tilman_biodiversity_2006; @franssen_annual_2011;
@grace_integrative_2016; @pennekamp_biodiversity_2018; @houlahan_negative_2018;
@olivier_urbanization_2020]. However, these studies mainly focused on communities
composed of a single trophic level, generally primary producers, thereby
ignoring potential effects of food-web structure [but see
@scherber_bottom-up_2010].
Conversely, theoretical works suggest that both species richness and food-web
structure affect ecosystem properties such as productivity, total biomass and
its temporal stability. For example, the presence of species at higher trophic
levels may increase total primary production through top-down effects, which
result in the selection of primary producers that are larger and more
complementary in their use of resources [@wang_biodiversity_2018]. In communities
including producers and consumers, the food-web connectance is predicted to
modulate the diversity-productivity relationship [@thebault_food-web_2003;
@thebault_cascading_2007]. Similarly, theoretical studies generally show strong
links between food-web structure and stability [e.g.  @may_will_1972;
@dunne_network_2006; @neutel_reconciling_2007; @duffy_why_2009;
@stouffer_compartmentalization_2011]. Overall, theoretical
models suggest that both species richness and food-web structure should affect
community biomass and its stability, but empirical evidence of such effects are
lacking.

While community stability is measured in many ways in theoretical studies, most
empirical studies measure community stability as the inverse of the coefficient
of variation of community productivity, biomass or abundance, also called
temporal stability [@kefi_advancing_2019]. This stability metric is also well
grounded in theory [e.g. @thebault_trophic_2005; @loreau_species_2008;
@thibaut_understanding_2013] and can be decomposed into two components:
population stability and synchrony [@thibaut_understanding_2013]. Studies
considering a single trophic level [@yachi_biodiversity_1999;
@tilman_biodiversity_2006; @loreau_species_2008; @olivier_urbanization_2020] as
well as theoretical prediction for multitrophic communities
[@thebault_trophic_2005] tend to highlight that asynchrony in population
dynamics is key for the stabilising effects of species richness on community
biomass. Theory predicts that higher food-web connectance may increase
interaction strength and thereby increase population variability,
counterbalancing the positive effect of species richness on the temporal
stability of community biomass [@thebault_trophic_2005]. Predators can also have
cascading effects on both population variability and synchrony at lower trophic
levels [@teng_dynamics_2004; @shanafelt_stability_2018]. Overall, these
theoretical studies show that the effects of species richness and food-web
structure, such as connectance and average trophic level, are potentially
conflicting or synergistic in such a way that they need to be simultaneously
assessed.

The joint study of the effects of species richness and food-web structure on
community biomass and its stability is important in the context of environmental
changes. Recent studies comparing natural communities showed that environmental
characteristics determine in part community biomass [@grace_integrative_2016;
@woods_testing_2020] and its temporal stability [@franssen_annual_2011;
@hansen_climate_2013; @bluthgen_land_2016; @de_boeck_patterns_2018;
@friedpetersen_drivers_2020]. The environment can affect ecosystem function or
stability directly. For instance, larger freshwater ecosystems are expected to
have more available resources and thus community biomass [@post_ecosystem_2000;
@doi_resource_2009] while agricultural intensification destabilises bird
communities by decreasing population stability [@olivier_urbanization_2020]. The
effects of environment can also be mediated by changes in species richness and
food-web structure. Larger ecosystems host food-webs with higher diversity and
trophic levels [e.g., @mchugh_dual_2010] and temperature is also known to affect
food-web diversity and structure  [e.g., @edeline_ecological_2013;
@ogorman_unexpected_2017; @gauzens_biodiversity_2020; @bonnaffe_trophic_2021].
Environmental characteristics have thus to be accounted for if we are to tease
apart the effects of species richness and food-web structure on community
biomass and its temporal stability.

Environment can affect functioning and stability directly or through effects on
resource availability, species richness or food-web
structure. Agricultural intensification and temperature changes can also...

---
#### Stream fish communities is a good model to study effects of horizontal & vertical diversity
---
Stream fish communities constitute a good model to study the effects of network
structure on community biomass and its temporal stability because their trophic
interactions are well-known. A substantial body of literature supports that
trophic interactions can be inferred through body size and diet information
[@jennings_weak_2001; @beckerman_foraging_2006; @gravel_inferring_2013;
@brose_predator_2019; @portalier_mechanics_2019]. Using long-term monitoring of stream fish
communities and inference of trophic interactions [@bonnaffe_trophic_2021], we
provide one of the first empirical tests of the expected relationship between
BEF and food-web complexity-stability accounting for environmental
characteristics.

@@

Fish communities were monitored across stream sections in metropolitan France
over the period 1995-2018 by the French Office of Water and Aquatic Ecosystems
(ONEMA) using electrofishing. Two different standardised protocols were used for
small and large streams (detailed in appendix B, part 1.1). In the case where a
stream was sampled alternatively with one of the two sampling protocols, we only
included data derived from the most frequently used sampling method. The
sampling season spanned from late spring to autumn. There was low heterogeneity
in the sampling month within stations, with the median sampling month being
mid-August and the median standard deviation of the sampling month was $0.7$ month,
i.e. three weeks.

We selected the stations that had been sampled ten years or more. As the concept
of temporal stability often referred to the variability around an equilibrium
point [@donohue_navigating_2016], we selected the stations that did not show temporal
trends in community biomass (detailed in appendix B, part 2). However, the
results were robust to the relaxation of this hypothesis (Fig. S7). These
selection steps resulted in 99 stations distributed over seven hydrographic
basins (Fig. \@ref(fig:matmet)A).

@@

The fishes collected by electrofishing were grouped in batches according to
their body size class and  species identity (Fig. S1, appendix B, part 1.1).
When individuals were too numerous in a given batch, the body size of a
subsample of the individuals was measured. The body size of the unmeasured
individuals was inferred under the assumption that the  body size of the
individuals in a batch follows a normal distribution (Fig. S1, appendix B,
part 1.2). The body mass of each individual was then estimated following the
results of two meta-analyses on fishes [@lleonart_removing_2000;
@froese_cube_2006], using an allometric law $B_i = 0.01 \times l_i^{3.03}$, with
$B$ the body mass in grams, $l$ the body size in millimetres and $i$ the
individual. Population biomass and community biomass at each sampling event were
then calculated by summing individual body masses. To account for the variation
in sampling effort among sites, population biomass and community biomass were
corrected by the surface sampled (range 
`r range_surface[1]` - `r range_surface[2]` square metres), and expressed in
grams per square metre ($gm^{-2}$). In the analysis, the community biomass of a
site was defined as the median of the community biomass across sampled years.

@@

Temporal biomass stability was defined as the inverse of biomass variability.
Variability was computed as Coefficient of Variation (CV) of the community
biomass over time ($CV_i = \sigma_i \mu_i^{-1}$, with $\sigma_i$ and $\mu_i$
being respectively the standard deviation and the temporal average of the annual
community biomass of the station $i$). Then, temporal stability ($S_i$) was
equal to $CV_i^{-1} = \mu_i \sigma_i^{-1}$. The variability of community
biomass ($CV_{com, i}$) can be decomposed into a synchrony ($\phi$) component
and a weighted average population variability ($\overline{CV_{sp, i}}$)
component according to @thibaut_understanding_2013:

@@

To assess the mechanisms driving stability of community biomass, we thus in addition measured
the synchrony and the weighted average population variability $\overline{CV_{sp,
i}}$, the average CV at the population-level in the station $i$, weighted by the
relative biomass of the species populations, $B_{k,i}$.

Synchrony of the station $i$, $\phi_i$, is the ratio between the variance of
community biomass over time ($\sigma^2_{x_{T,i}}$) and the squared sum of the
standard deviation of population biomass over time ($\sigma_{x_{k,i}}$), $i$
being the station $i$ and $k$ the species $k$ [@loreau_species_2008]. Synchrony
varies from 0 to 1; 0 indicating complete asynchrony and 1 perfect synchrony.
Consequently, the lower is synchrony, the higher are the compensatory dynamics.

@@

Species richness in a given site was defined as the total number of species
present across all the sampling events. We controlled species richness for
sampling effort by dividing the total number of species by the total surface
sampled in the given station. Species richness was then expressed in number of
species per square metre sampled.

### Food-web inference

@@

We determined the food-web structure in two steps [@bonnaffe_trophic_2021]: (1) we
built a metaweb describing the trophic interactions among all possible fish size
classes and species as well as with non-fish resource classes (Fig. S2), then
(2) we determined the food-web corresponding to each sampling event by
extracting from the metaweb the subset of fish species and size classes present
at a given site and sampling event (Fig. \@ref(fig:matmet)C and D).

The food-web was inferred from body size and ontogenic diet shift following
previous studies [@gravel_inferring_2013; @poisot_structure_2016;
@brose_predator_2019; @bonnaffe_trophic_2021]. The diet of stream fishes, as for
many organisms, shifts throughout an individual’s life, and thereby with body
size. For example, larvae and small juveniles feed on plankton while adults may
feed on algae and fishes, with bigger individuals eating bigger fishes. Each
fish species was divided into nine body size classes, each corresponding to a
trophic species (appendix B, Fig. S2). A trophic species constitutes a group of
individuals from the same species and size class, which are expected to share
the same sets of prey and predators. Additionally, we added seven resource nodes
present in the ontogenic food diet database: detritus, biofilm, phytoplankton,
zooplankton, macrophages, phytobenthos and zoobenthos (grey nodes in Fig.
\@ref(fig:matmet)C & D). The trophic species and the resources constituted the
412 nodes of the metaweb, i.e. 45 fish species divided into nine body size
classes plus seven resource nodes. The inference of trophic interactions was
shown to be more accurate when considering body size classes, i.e. trophic
species, than species [@jennings_weak_2001].

The feeding interactions of a trophic species were determined by both its
species identity and its body size. The ontogenic fish diet database was filled
according to fishbase [@froese_fishbase_2021] and literature (appendix A, Table
S2). Each fish species had two or three life stages which are delimited by its
body size range (appendix B, Fig. S2). The life stage of a fish trophic species
was determined by the centre of its body size range, hereafter midpoint
(appendix B, Fig. S2).

For a piscivorous trophic species, the fish-fish trophic links were defined in
two steps and based on the body size ratio between predators and prey. First of
all, the predation window of a piscivorous trophic species was defined as 3% to
45% of its midpoint [@mittelbach_ontogeny_1998; @claessen_impact_2002;
@hart_handbook_2008]. Then, a trophic link was set between the piscivorous trophic
species and every trophic species whose midpoint in body size range was included
in the predation window (appendix B, Fig. S2). The fish-resource trophic links
were set between the trophic species and the resources nodes according to the
food items of their size class. Lastly, the trophic links among resource nodes
were set according to the literature [@allan_stream_2007; @hart_handbook_2008].

### Food-web structure

For each food-web, corresponding to a particular site and year, we computed the
connectance and mean trophic level. Food-web connectance describes the level of
generalism in the network and is calculated as the actual number of interactions
divided by the number of interactions if all species interact ($C = L / N^2$, L
and N being resp. the number of links and of nodes). The trophic level of the
nodes was computed as one added to the average trophic level of its food items.
The average trophic level $\overline{T}$ was computed as the average trophic
level of the trophic species weighted by their biomass (excluding resource nodes
for which we had no biomass information), so $\overline{T} = \sum_i^N{B_iT_i
\times (\sum_i^N{B_i})^{-1}}$, with $T_i$ and $B_i$ being respectively the
trophic level and the biomass of the node $i$. For each site, the median of
connectance and average trophic level over time was used in subsequent analysis
(Fig. \@ref(fig:matmet)C & D).

@@

We characterised the sites by their altitude, slope, distance from source,
stream width, depth and strahler order, i.e. its position in the stream network
hierarchy. We obtained data on Biological Oxygen Demand (BOD), which is often
related to enrichment and eutrophication in aquatic systems
[@mallin_factors_2006], water temperature, flow and from the naiades database
[(www.naiades.eaufrance.fr)](http://www.naiades.eaufrance.fr/). As data from naiades are not collected on the same
site as our fish sampling sites, we performed an interpolation of the water
temperature and flow based on the spatial stream network models
[@isaak_applications_2014; @hoef_ssn:_2014]. The interpolation procedure is
detailed in appendix B (part 1.4).

We had one environmental variable value for each sampling event, except for
water temperature for which records began in 2006. Then, we computed the average
and temporal CV values of water temperature, flow, BOD, river width and depth.
As habitat and environmental variables are often collinear, a principal
component analysis (PCA) was performed on the 14 scaled variables. We kept in
the analysis the two first principal components based on the elbow method on the
eigenvalues distribution (appendix B, Fig. S5). The two first principal
components explained `r tot_var_expl_pca`% of the variance (Fig.
\@ref(fig:matmet)B). A varimax axis rotation was performed to maximise the
representation of the environmental variables by the principal components
[@kaiser_varimax_1958; @revelle_psych_2019]. The first axis represented a
gradient from small to big streams, characterised by their width and depth, but
also a gradient of distance from source. We reversed the values
of the second axis, so that is represented a gradient of increasing average
water temperature, average BOD and decreasing altitude and slope. Further
description of the variables are provided in appendix B (part 1.4).

@@

We analysed the relationships among environment, species richness, food-web
structure, community biomass and its temporal stability with structural equation
models as they enable the assessment of potential direct and indirect
relationships among multiple variables [@grace_structural_2008;
@grace_causal_2014; @lefcheck_piecewisesem_2016; @grace_integrative_2016]. The
structural equation model explains the median annual community biomass as a
function of the environment, species richness and food-web structure
[@maureaud_biodiversityecosystem_2019; @woods_testing_2020]. We assumed that the
environment, species richness and food-web structure directly affect community
biomass (Fig. \@ref(fig:sem-bm)A), that species richness affects food-web
structure [@winemiller_must_1989; @dunne_network_2006; @thebault_stability_2010]
and that environment affects species richness and food-web structure.

We used the same model structure to assess the drivers of temporal stability of
community biomass (Fig. \@ref(fig:sem-stab)A). Instead of the annual community
biomass, we modelled the synchrony and population variability which in turn
fully determined the temporal stability of community biomass. We linearized the
mathematical relationship linking temporal stability of community biomass $S_i$
to $\overline{CV_{sp, i}}$ and $\phi$ by taking the log of temporal stability
such as

@@

The structural equation models were based on Gaussian linear mixed-effects
models. All variables except the connectance and average trophic level were
log-transformed to fulfil model requirements (see residual plots and data
transformation in Fig. S5 and S6, Table S5, appendix B). We set the seven
hydrographic basins as a random effect on the intercept to account for their
heterogeneity. We checked for the absence of multicollinearity with Variance
Inflation Factor (VIF, Table S6, appendix B). The slopes estimated by the
structural equation models and reported in the main text were standardised in
order to compare their magnitude ($r_\delta = \beta \frac{\sigma_x}{\sigma_y}$,
$\beta$ being the unstandardised slope, $\sigma_x$ and $\sigma_y$ being
respectively the standard deviation of the predictor variable $x$ and of the
response variable $y$). Therefore, the standardised coefficients expressed the
variation of $x$ and $y$ in standard deviation units. According to the equation
\@ref(eq:stab), the unstandardised coefficients of synchrony and population
variability determining
temporal stability of community biomass are respectively -0.5 and -1.0. When
standardised, the relative comparison of those coefficients comes down to a
comparison of the standard deviation of these two variables. The deviation from
unstandardised effect values (i.e. -0.5 and -1.0) indicates that one component of community stability
displays more variations than the other, and that the determinants of
the more variable component will be those affecting community
stability the most. The paths whose slope had a $P$ value inferior or equal to $0.05$
were reported in the main text, all the path values are presented in Table S7
and S8. Indirect effects on community biomass and its temporal stability were
computed by multiplying slopes along the path. Only significant direct slopes
were included in the calculation of indirect effects. Total effects were
computed as the sum of direct and indirect effects.

Environmental variables were interpolated with the `openStars` and `SSN` R
packages [@kattwinkel_openstars_2018; @hoef_ssn:_2014]. Linear models and SEMs
were computed with respectively the nlme and piecewieseSEM R packages
[@pinheiro_nlme_2020; @lefcheck_piecewisesem_2016]. The code used for the
analysis and to generate the manuscript are accessible on github
(alaindanet/fishcom). The analyses were performed with R version 3.6.3
[@r_core_team_r_2020].

@@
